---
title: "p8105_hw2_qz2493.Rmd"
author: "Qingyue Zhuo qz2493"
date: "2022-10-03"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

# load the packages
```{r}
library(tidyverse)
library(readxl)
library(dplyr)
library(lubridate)
```

# Problem 1
```{r}
subway_data = 
  read_csv(
    "data/Subway.csv", 
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>%
  janitor::clean_names() %>%
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>%
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

The dataset contains the following variables:`r names(subway_data)`. I used janitor::clean_names to do the basic cleaning. The dimension of the dataset is `r nrow(subway_data)`*`r ncol(subway_data)`. These data are not "tidy": route number should be a variable, as should route..

```{r}
subway_distinct = distinct(subway_data, line, station_name, .keep_all = TRUE)
```

There are `r nrow(subway_distinct)` stations. 

```{r}
ada_compliant = filter(subway_distinct, ada == 'TRUE')
entry_without_vending = 
  filter(subway_distinct, vending == "NO") %>%
  pull(entry) %>% 
  mean
```

`r nrow(ada_compliant)` stations are ADA. `r 100*(nrow(entry_without_vending)/nrow(subway_distinct))`% of station entrances without vending allow entrance.

Lastly, we write a code chunk to identify stations that serve the A train, and to assess how many of these are ADA compliant. As a first step, we tidy the data as alluded to previously; that is, we convert `route` from wide to long format. After this step, we can use tools from previous parts of the question (filtering to focus on the A train, and on ADA compliance; selecting and using `distinct` to obtain dataframes with the required stations in rows).
```{r}
subway_data %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

subway_data %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```


# Problem 2
# Read and clean the Mr. Trash Wheel sheet
```{r}
trash_data = 
  read_excel("data/trash.xlsx", sheet = 1, range = cell_cols("A:N"),  skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  drop_na(month) %>%
  mutate(
    sports_balls = as.integer(round(sports_balls))
  )
trash_data
```

# Read and clean the Professor Trash Wheel sheet
```{r}
professor_data = 
  read_excel("data/trash.xlsx", sheet = 2, range = cell_cols("A:M")) %>%
  janitor::clean_names() %>%
  drop_na() 

professor_data
```

#Combine two dataset
```{r}
combine_data = 
  left_join(trash_data, professor_data, by = "date")
combine_data
```

## The trash_wheel data set has `r nrow(trash_data)` observations of `r ncol(trash_data)` variables. The key variables are `r names(trash_data)`. The total number of sports balls collected by Mr. Trash Wheel in 2020 is `r sum(pull(trash_data,sports_balls, year == 2020))` `r sum(pull(filter(trash_data, year == 2020), sports_balls))`.

## The professor trash wheel data set has `r nrow(professor_data)` observations of `r ncol(professor_data)` variables. The key variables are `r names(professor_data)`. The total weight of trash collected by Professor Trash Wheel is `r sum(pull(professor_data, weight_tons))` tons.

## The combined data set has `r nrow(combine_data)` observations of `r ncol(combine_data)` variables. 

# Problem 3-president taking values gop and dem
# clean and organize pols-month dataset
```{r}
pols_data = 
  read_csv("data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year","month","day"), sep = "-") %>%
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    day = as.integer(day),
    month = month.name[month]) %>%
  select(-day) %>%
  relocate(year, month, prez_gop, prez_dem) %>%
  mutate(
    prez_gop = na_if(prez_gop, 0),
    prez_dem = na_if(prez_dem, 0)) %>%
  pivot_longer(
    prez_gop:prez_dem,
    names_to = "name",
    names_prefix = "prez_",
    values_to = "president") %>%
  drop_na() %>%
  select(-president) %>%
  rename(president = name) %>%
  select(year, month, president, everything())

pols_data
```

#clean and organize snp data set
```{r}
snp_data = 
  read_csv("data/snp.csv") %>%
  janitor::clean_names() %>%
  mutate(
    month = month(mdy(date)),
    year = year(mdy(date)),
    year = ifelse(year >= 2050, year - 100, year)) %>%
  select(year, month, close) %>%
  mutate(
    month = month.name[month]) %>%
  arrange(year, month)

snp_data
 
```

# clean and organize unemployment data
```{r}
unemploy_data = 
  read_csv("data/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = 'month',
    values_to = 'unemployment') %>%
  mutate(
    month = recode(month, "jan" = 1, "feb" = 2, "mar" = 3, "apr" = 4, "may" = 5, "jun" = 6, "jul" = 7, "aug" = 8, "sep" = 9, "oct" = 10, "nov" = 11, "dec" = 12),
    month = month.name[month]
  )

unemploy_data
```

#combine the data set
```{r}
pols_snp =
  left_join(pols_data, snp_data, by = c("year", "month"))
pols_snp_unemploy = 
  left_join(pols_snp, unemploy_data, by = c("year", "month")) %>%
  arrange(year, month)

pols_snp_unemploy
```

Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).

## Pols-month dataset has `r nrow(pols_data)` observtions of `r ncol(pols_data)` variables. It contains the number of presidents who are democratic or republican at certain given time. For example, at a given month and year, whether the president belongs to democratic or republican party, and the number of republican or democratic governors, senators and representatives.

## Snp dataset has `r nrow(snp_data)` observations of `r ncol(pols_data)` variables. It introduces the Standard & Poor’s stock market index, which shows the closing values of the S&P stock index on the associate date.

## Unemployment dataset has `r nrow(unemploy_data)` observations of `r ncol(unemploy_data)` variables. It reveals the percentage of unemployment at a given month and year.

## The resulting data set has `r nrow(pols_snp_unemploy)` observations of `r ncol(pols_snp_unemploy)` variables. Year range is 1947 to 2015, and the key variables are `r names(pols_snp_unemploy)`. 
