---
title: "p8105_hw2_qz2493.Rmd"
author: "Qingyue Zhuo qz2493"
date: "2022-10-03"
output: html_document
---
# load the packages
```{r}
library(tidyverse)
library(dplyr)
```

# Problem 1- include ada notes or not?- are these data tidy?-how to reformat the data?
```{r}
subway_data = 
  read_csv("data/Subway.csv") %>%
  janitor::clean_names() %>%
  select(line:entry, vending, ada) %>%
  mutate(
    entry = recode(entry, 'YES' = '1', 'NO' = '0'))

subway_distinct = distinct(subway_data, line, station_name, .keep_all = TRUE)
ada_compliant = filter(subway_distinct, ada == 'TRUE')
entry_without_vending = filter(subway_distinct, entry == "1", vending == "NO")


```
The dataset contains the following variables:`r names(subway_data)`. I used janitor::clean_names to do the basic cleaning. The dimension of the dataset is `r nrow(subway_data)`*`r ncol(subway_data)`. These data are tidy.

There are `r nrow(subway_distinct)` stations. `r nrow(ada_compliant)` stations are ADA. `r 100*(nrow(entry_without_vending)/nrow(subway_distinct))`% of station entrances without vending allow entrance.

# Problem 2
# Problem 3

